---
title: "Pollack in ICES Divisions 89a"
author: "Paz Sampedro"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Background information

Pollack.27.89a is a stock in ICES Category 5. Currently, the management advice is based on commercial landings. 

Pollack is a bentopelagic species. Main commercial fleets included gillnetters and longliners from France (80% landings) and gillnetters and longliners from Spain (17% landings). Recreational landings are supposed to be high (at the same level as commercial landings, they have been estimated. 

## Available data 
1. Annual commercial landings (in tonnes): 1986-2018. 
2. Abundance Index  "France gillneters mesh > 90 mm, operating in div 8a, 2nd semester" (tonnes/fishing sequence): 2005-2018.

## Fitting SPICT model

### Starting session and download libraries 
```{r results='hide', message=FALSE, warning=FALSE}

# Package from github:
  #devtools::install_github("mawp/spict/spict")
  #library(remotes)
  #install_github("DTUAqua/spict/spict") 

rm(list=ls(all=TRUE))

# Download libraries
library(spict)
library(icesAdvice)
library(ellipse)
library(formatR)
```

### Creating data object

```{r results='hide', message=FALSE, warning=FALSE}

# Catch data: Only Commercial Landings. *Year 1999: mean of 1998 and 2000 as French landings were missing

pol89aC <- data.frame(obsC = c(2806,2918,2582,1973,1900,2168,1958,1513,1955,1679,1354,1378,1165,1322,1479,1746,1972,1663,1726,1986,2126,1847,2313,1812,1682,2032,1520,1811,1959,1610,1661,1481,1512),timeC = 1986:2018)

# Index data: Commercial index FR-GNS>90mm-8a-2s (tonnes/fishing sequence)
pol89aI <- data.frame(obsI = c(0.1151,0.0663,0.1292,0.1289,0.1244,0.1079,0.1966,0.1743,0.1565,0.1500,0.1438,0.1205,0.1222,0.1117),timeI = c(2005:2018)+0.5)

# Create a list for input data
pol89a <- list(timeC = pol89aC$timeC, obsC = pol89aC$obsC,timeI = list(pol89aI$timeI),obsI = list(pol89aI$obsI))

# Check list 
pol89a <- check.inp(pol89a)
```

Plot Raw data:

```{r Figure1, message=FALSE, warning=FALSE, fig.cap = "Raw data ", fig.align = "center"}
plotspict.data(pol89a)
```

To highlight:

- Catch and abundance series are overlapped only for a short period of time (14 years).

- There is not contrast in catch series during 2005-2018.


Plot Fitting linear regression:
```{r Figure2, message=FALSE, warning=FALSE, fig.cap = "Fitting linear regression", fig.align = "center"}
plotspict.ci(pol89a)
```


### Model Fitting 

#### Run 1. Default priors and all parameters estimated
 Not converged. Model did not obtain proper convergence!

```{r results='hide', message=FALSE, warning=FALSE}
res_pol89aDefault <- fit.spict(pol89a)
```

#### Run 2. Shorten Catch series to 2005-2018, overlapping with abundance index.

Not converged. Shorten series gave totally unrealistic estimates, testing different configurations:

```{r results='hide', message=FALSE, warning=FALSE}
pol89aShorten = shorten.inp(pol89a, 2005, 2018)
res_pol89aShorten <- fit.spict(pol89aShorten)
```
#### Run 3. Fix parameter to Schaefer production curve (initial parameter). Symmetric productive curve (BMSY/K=0.5)

No converged:
```{r message=FALSE, warning=FALSE}
pol89aSchaefer <- pol89a
pol89aSchaefer$ini$logn <- log(2)
pol89aSchaefer$phases$logn <- -1
res_pol89aSchaefer <- fit.spict(pol89aSchaefer)
res_pol89aSchaefer 
```

#### Run 4. Set priors for the ratio between biomass in the initial year relative to K, mean of log(0.5) and sd of 0.2

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
pol89aBkfrac <- pol89a
pol89aBkfrac$priors$logbkfrac <- c(log(0.5),0.2,1)
res_pol89aBkfrac <- fit.spict(pol89aBkfrac)
res_pol89aBkfrac

# Summary of estimates
round(sumspict.parest(res_pol89aBkfrac),2)

# Reference points
sumspict.drefpoints(res_pol89aBkfrac) 

# Basic plotting of the results 
plot(res_pol89aBkfrac)

# Convergence checks

# Convergence
res_pol89aBkfrac$opt$convergence 

# All the sd were estimated
all(is.finite(res_pol89aBkfrac$sd))

# No violation of assumptions: bias, correlation and normality:
res_pol89aBkfrac <- calc.osa.resid(res_pol89aBkfrac)
plotspict.diagnostic(res_pol89aBkfrac)
```

#### Retrospective: no trends and all runs inside the confidence intervals

Problems of convergence:

```{r message=FALSE, warning=FALSE}
rep<-retro(res_pol89aBkfrac, nretroyear=5)

```

#### Run 5:

- Fix parameter to Schaefer production curve (initial parameter). Symmetric productive curve (BMSY/K=0.5) 

- Set priors for the ratio between biomass in the initial year relative to K, mean of log(0.5) and sd of 0.2

```{r message=FALSE, warning=FALSE}
pol89aSchaefBkfrac <- pol89a
# Set priors for B/K
pol89aSchaefBkfrac$priors$logbkfrac <- c(log(0.5),0.2,1)
# Fixing n at 2
pol89aSchaefBkfrac$ini$logn <- log(2)
pol89aSchaefBkfrac$phases$logn <- -1

# Fit the model
res_pol89aSchaefBkfrac <- fit.spict(pol89aSchaefBkfrac)
# Results 
res_pol89aSchaefBkfrac

# Summary of estimates
round(sumspict.parest(res_pol89aSchaefBkfrac),2)

# Reference points
sumspict.drefpoints(res_pol89aSchaefBkfrac) 

# Basic plotting of the results 
plot(res_pol89aSchaefBkfrac)

# Convergence checks

# Convergence
res_pol89aSchaefBkfrac$opt$convergence 

# All the sd were estimated
all(is.finite(res_pol89aSchaefBkfrac$sd))

# No violation of assumptions: bias, correlation and normality:
res_pol89aSchaefBkfrac <- calc.osa.resid(res_pol89aSchaefBkfrac)
plotspict.diagnostic(res_pol89aSchaefBkfrac)
```

#### Retrospective: no trends and all runs inside the confidence intervals

All runs converge and estimates are inside the confidence intervals.
\
Problem with last year?. Trend changed relative to previous runs.

```{r message=FALSE, warning=FALSE}
rep <- retro(res_pol89aSchaefBkfrac, nretroyear=5)
plotspict.retro(rep)
```

#### Checking robustness to initial parameter values

All runs converged. It seems that there are two optima (see $resmat). 18 runs with the same realistic estimates; but 12 converged models with unrealistic estimates.

```{r message=FALSE, warning=FALSE}
set.seed(123)
check.ini(pol89aSchaefBkfrac, ntrials=30)
```


# Questions for Workshop - September2020:

Run 5, that fixes n of productive curve (BMSY/K= 0.5) and uses priors for B1/K, converged well and the output estimates are realistic. However, the model is sensitive to initial values and the confidence intervals are quite wide.

1. How the robustness can be improved?. To address the sensitivity to initial values.
2. Acting (fixing, set priors) on alfa and/or beta could help to narrow the confidence intervals?




