---
title: "Octopus spict"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

### Run SPiCT model for Octopus vulgaris in the Galician coast

```{r}

library(tidyverse)
library(spict)
library(ellipse)

xtab<-function (x,caption='Table X.',
                file=stdout(),width='"100%"',
                cornername='',
                dec=rep(1,ncol(x))) {
  nc<-ncol(x)
  lin<-paste('<table width=',width,'>',sep='')
  lin<-c(lin,sub('$','</td></tr>',
                 sub('\\. |\\.$','.</b> ',
                     sub('^',paste('<tr><td colspan=',nc+1,'><b>',sep=''),caption))))
  hr<-paste('<tr><td colspan=',nc+1,'><hr noshade></td></tr>',sep='')
  lin<-c(lin,hr)
  cnames<-colnames(x)
  cnames<-paste(sub('$','</b></td>',sub('^','<td align=right><b>',cnames)),collapse='\t')
  lin<-c(lin,paste('<tr>',paste('<td align=left><b>',cornername,'</b></td>',sep=''),cnames,'</tr>'))
  lin<-c(lin,hr)
  rnames<-sub('$','</b></td>',sub('^','<tr> <td align=left><b>',rownames(x)))
  #x<-sapply(1:ncol(x),function(i)sub('NA','  ',format(round(x[,i],dec[i]))))
  x<-sapply(1:ncol(x),function (i) sub('NA','  ',
                                       formatC(round(x[,i],dec[i]),digits=dec[i],format='f')))
  for (i in 1:nrow(x)) {
    thisline<-paste(rnames[i],paste(sub('$','</td>',
                                        sub('^','<td align=right>',x[i,])),
                                    collapse='\t'),'</tr>', sep='')
    lin<-c(lin,thisline)
  }
  lin<-c(lin,hr)
  lin<-c(lin,'</table><br>\n')
  writeLines(lin,con=file)
}

```

#1# Read data


```{r}

octopus<-read.table(file="/Users/jaimeoterovillar/Documents/IEO_work/IMPRESS/Spict-IMPRESS/Octopus/pulpo_catch_cpue.txt",header=T,dec=".",sep=",")

head(octopus)


octopus.8c<-filter(octopus,Zone_Ices=="Ices_8.c") # Data sets for Ices 8c
octopus.9a<-filter(octopus,Zone_Ices=="Ices_9.a") # Data sets for Ices 9a

```


#2# SPiCT for 8c (North) -------------------------------------------------------

#2.1# Set up data and check

```{r}


inp8c<-list(timeC = octopus.8c$FishingSeason,
            obsC  = octopus.8c$Catches, # in Tons
            obsI  = octopus.8c$Mean/1000, # Standardized biomass per 50 traps
            timeI = octopus.8c$FishingSeason)

inp8c<-check.inp(inp8c)
inp8c$dtc

```


#2.2# Plot data

```{r}

# plotspict.data(inp8c) # simple plotting
plotspict.ci(inp8c) # advanced plotting

```


#2.3# Fitting the model and get summaries

```{r}

res8c<-fit.spict(inp8c)

names(res8c) # contents of the model

capture.output(summary(res8c)) # model summary (all information)

round(sumspict.parest(res8c),2) # model parameter estimates with 95% CI
sumspict.drefpoints(res8c) # deterministic reference points (Drp)
sumspict.srefpoints(res8c) # stochastic reference points (Srp)
sumspict.states(res8c) # states with 95% CI for the last year of the series
sumspict.predictions(res8c) # predictions with 95% CI for the next year

```


#2.4# Plotting results

```{r}

plot(res8c) # all plots together

par(mfcol=c(2,3))
plotspict.biomass(res8c,qlegend=F,ylim=c(0,4000)) # biomass plot
plotspict.bbmsy(res8c,qlegend=F,ylim=c(0,3)) # relative biomass plot
plotspict.f(res8c,qlegend=F,rel.axes=F,rel.ci=F) # fishing mortality plot
plotspict.ffmsy(res8c,qlegend=F) # relative fishing mortality plot
plotspict.catch(res8c,qlegend=F,ylim=c(0,2000)) # catch plot
plotspict.fb(res8c)#,ylim=c(0.2,0.4),xlim=c(0,4000)) # kobe plot

```


#2.5# Model residuals and diagnostics

```{r}

res8c$opt$convergence # ok
all(is.finite(res8c$sd)) # ok

res8c.res<-calc.osa.resid(res8c) # get residuals

plotspict.diagnostic(res8c.res) # plot residuals

```


#2.6# Get important quantities

```{r}

list.quantities(res8c) # all quantities from the model

param.8c<-sumspict.parest(res8c)
refpoints.det.8c<-sumspict.drefpoints(res8c)
refpoints.stoch.8c<-sumspict.srefpoints(res8c)
sate.8c<-sumspict.states(res8c)
pred.8c<-sumspict.predictions(res8c)

xtab(param.8c,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/Params8c.html",dec=rep(4,ncol(param.8c)))

xtab(refpoints.det.8c,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/RefPointsDet8c.html",dec=rep(4,ncol(refpoints.det.8c)))

xtab(refpoints.stoch.8c,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/RefPointsSto8c.html",dec=rep(4,ncol(refpoints.stoch.8c)))

xtab(sate.8c,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/State8c.html",dec=rep(4,ncol(sate.8c)))

xtab(pred.8c,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/Forecast8c.html",dec=rep(4,ncol(pred.8c)))

```


#2.7# Restrospective analysis

```{r}

ret.8c<-retro(res8c,nretroyear=3)

plotspict.retro(ret.8c)

```



#3# SPiCT for 9a (South) -------------------------------------------------------

#3.1# Set up data and check

```{r}

inp9a<-list(timeC = octopus.9a$FishingSeason,
            obsC  = octopus.9a$Catches, # in Tons
            obsI  = octopus.9a$Mean/1000, # in Tons
            timeI = octopus.9a$FishingSeason)

inp9a<-check.inp(inp9a)
inp9a$dtc

```


#3.2# Plot data

```{r}

# plotspict.data(inp9a) # simple plotting
plotspict.ci(inp9a) # advanced plotting

```


#3.3# Fitting the model and get summaries

```{r}

res9a<-fit.spict(inp9a)

names(res9a) # contents of the model

capture.output(summary(res9a)) # model summary (all information)

round(sumspict.parest(res9a),2) # model parameter estimates with 95% CI
sumspict.drefpoints(res9a) # deterministic reference points (Drp)
sumspict.srefpoints(res9a) # stochastic reference points (Srp)
sumspict.states(res9a) # states with 95% CI for the last year of the series
sumspict.predictions(res9a) # predictions with 95% CI for the next year

```


#3.4# Plotting results

```{r}

plot(res9a) # all plots together

par(mfcol=c(2,3))
plotspict.biomass(res9a,qlegend=F,ylim=c(0,4000)) # biomass plot
plotspict.bbmsy(res9a,qlegend=F,ylim=c(0,4)) # relative biomass plot
plotspict.f(res9a,qlegend=F,rel.axes=F,rel.ci=F) # fishing mortality plot
plotspict.ffmsy(res9a,qlegend=F) # relative fishing mortality plot
plotspict.catch(res9a,qlegend=F,ylim=c(0,3000)) # catch plot
plotspict.fb(res9a)#,ylim=c(0.2,0.4),xlim=c(0,4000)) # kobe plot

```


#3.5# Model residuals and diagnostics

```{r}

res9a$opt$convergence # ok
all(is.finite(res9a$sd)) # ok

res9a.res<-calc.osa.resid(res9a) # get residuals

plotspict.diagnostic(res9a.res) # plot residuals

```


#3.6# Get important quantities

```{r}

list.quantities(res9a) # all quantities from the model

param.9a<-sumspict.parest(res9a)
refpoints.det.9a<-sumspict.drefpoints(res9a)
refpoints.stoch.9a<-sumspict.srefpoints(res9a)
sate.9a<-sumspict.states(res9a)
pred.9a<-sumspict.predictions(res9a)

xtab(param.9a,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/Params9a.html",dec=rep(4,ncol(param.9a)))

xtab(refpoints.det.9a,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/RefPointsDet9a.html",dec=rep(4,ncol(refpoints.det.9a)))

xtab(refpoints.stoch.9a,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/RefPointsSto9a.html",dec=rep(4,ncol(refpoints.stoch.9a)))

xtab(sate.9a,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/State9a.html",dec=rep(4,ncol(sate.9a)))

xtab(pred.9a,caption="Forecast",cornername="",
     file="/Users/jaimeoterovillar/Desktop/Forecast9a.html",dec=rep(4,ncol(pred.9a)))

```


#3.7# Restrospective analysis

```{r}

ret.9a<-retro(res9a,nretroyear=3)

plotspict.retro(ret.9a)

```
