---
title: "SPiCTing with Strait of Gibraltar blackspot seabream target fishery data (Spain and Morocco) - Exercise 2"
author: "J. Gil"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data used
A short time-series of species catch and biomass index (standardized Spanish CPUE) by quarter (2009-2018)

##Loading required Packages
```{r, eval=TRUE}
require(spict)
require(gdata)
```

##Required data for Ex.2
Landings came from Spain and Morocco are in tons by quarter (2009-2018)
Biomass index: standardized CPUE (k/h)from Spain VMS data by quarter (2009-2018) 
```{r, eval=TRUE}
dir()
dat2 <- read.xls("sbr_quarterlydata.xls")
```

##"inp" for the SPiCT model. 
```{r, eval=TRUE}
inp3 <- list(timeC = dat2[,1],
            obsC = dat2[,2],
            timeI = dat2[,1]+0.125, 
            obsI = dat2[,3])
inp3 <- check.inp(inp3)
```

##Plotting data inputs
```{r}
plotspict.data(inp3)
plotspict.ci(inp3) ##Advanced plots
```

##Fitting the model
```{r}
fit3 <- fit.spict(inp3)
```

##Summary
```{r}
capture.output(summary(fit3))
```

#Plotting outputs
```{r}
plot(fit3)
```
Really uncertain but converge...short data series :-(

##Alternate model: fixing to Schaefer model
```{r}
inp4 <- inp3
inp4$phases$logn <- -1
inp4$ini$logn <- log(2)
inp4$priors$logn <- c(0,0,0)

fit4 <- fit.spict(inp4)

plot(fit4)

capture.output(summary(fit4))
```

Sometimes you required change the type of the reference point to deterministic, because there´s not biomass relative to Bmsy (NA) Figure 
```{r}
inp4$msytype <- "d"

fit4 <- fit.spict(inp4)

plot(fit4)

capture.output(summary(fit4))
```
More better...less uncertain :-)

# Checklist for the acceptance of a SPiCT assessment 

- 1: Convergence 
```{r}
fit3$opt$convergence 
fit4$opt$convergence 
```
Both models converge :-)

- 2: All variance parameters are finite 
```{r}
all(is.finite(fit3$sd))  
all(is.finite(fit4$sd))  
```
TRUE in both models :-)

- 3: No violation of model assumptions 
```{r}
res3 <- calc.osa.resid(fit3)
plotspict.diagnostic(res3)
res4 <- calc.osa.resid(fit4)
plotspict.diagnostic(res4)
```
Shapiro p-value significant value for Index in model 3 (normality of the residuals)!!
Shapiro p-value significant value for catches  (normality of the residuals) and also 
Bias p-value significant for Index in model 4 ( mean of the residuals is different from zero)!!

- 4: Consistent patterns in the Retrospective Analysis
```{r}
rep3=retro(fit3, nretroyear=3) #quite short data series (3 years instead 5)
plotspict.retro(rep3)
rep4=retro(fit4, nretroyear=3) #quite short data series (3 years instead 5)
plotspict.retro(rep4)
```
Quite weird, especially model 3 (with reference points not deterministic) :-(

- 5. Realistic production curve
```{r}
calc.bmsyk(fit3)
calc.bmsyk(fit4) # ALTERNATE Schaefer fix (0.5)
```
0.2783732 in model 3 :-(

- 6. High assessment uncertainty can indicate a lack of contrast in the input data or violation of
the ecological model assumptions
```{r}
calc.om(fit3)
calc.om(fit4)
```
B/Bmsy and F/Fmsy confidence intervals with order magnitude >1 in model 3 :-( 
B/Bmsy and F/Fmsy confidence intervals with order magnitude =0 in model 4 :-)

- 7. Initial values do not influence the parameter estimates
```{r}
check.ini(inp3, ntrials=33)
check.ini(inp4, ntrials=33)
```
All 33 trials converge in model 3 but distance in $resmat is  huge ina a couple of trials: 4224.10 and 10715.76!! :-(
All 33 trials converge in model 4 and  distance in $resmat is 0 in all casews :-)

# Extract values in TABLE format
Preparing TABLE format
```{r, include=FALSE}
xtab<-function(x,caption='Table X.', file=stdout(), width='"100%"', cornername='', dec=rep(1,ncol(x))){
  nc<-ncol(x)
  lin<-paste('<table width=',width,'>', sep='')
  lin<-c(lin,sub('$','</td></tr>',sub('\\. |\\.$','.</b> ',
                                      sub('^', paste('<tr><td colspan=',nc+1,'><b>',sep=''), caption))))
  hr<-paste('<tr><td colspan=',nc+1,'><hr noshade></td></tr>', sep='')
  lin<-c(lin,hr)
  cnames<-colnames(x)
  cnames<-paste(sub('$','</b></td>',sub('^','<td align=right><b>',cnames)), collapse='\t')
  lin<-c(lin,paste('<tr>',paste('<td align=left><b>',cornername,'</b></td>',sep=''),cnames,'</tr>'))
  lin<-c(lin,hr)
  rnames<-sub('$','</b></td>',sub('^','<tr> <td align=left><b>',rownames(x)))
  #x<-sapply(1:ncol(x),function(i)sub('NA','  ',format(round(x[,i],dec[i]))))
  x<-sapply(1:ncol(x),function(i)sub('NA','  ',formatC(round(x[,i],dec[i]),digits=dec[i], format='f')))
  for(i in 1:nrow(x)){
    thisline<-paste(rnames[i],paste(sub('$','</td>',sub('^','<td align=right>',x[i,])), collapse='\t'),'</tr>', sep='')
    lin<-c(lin,thisline)
  }
  lin<-c(lin,hr)
  lin<-c(lin,'</table><br>\n')
  writeLines(lin,con=file)
}
```

Then choose the model and its values can be extracted with these tables:
```{r}
Tab1 <- sumspict.parest(fit3); Tab1
Tab2 <- sumspict.srefpoints(fit3);Tab2
Tab3 <- sumspict.states(fit3);Tab3
Tab4 <- sumspict.predictions(fit3);Tab4
Tab1_Sfix <- sumspict.parest(fit4); Tab1_Sfix
Tab2_Sfix <- sumspict.srefpoints(fit4);Tab2_Sfix
Tab3_Sfix <- sumspict.states(fit4);Tab3_Sfix
Tab4_Sfix <- sumspict.predictions(fit4);Tab4_Sfix
```

