---
title: "SPiCTing with Strait of Gibraltar blackspot seabream target fishery data (Spain and Morocco) - Exercise 2+"
author: "J. Gil"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data used
A short time-series of species catch and biomass index (standardized Spanish CPUE) by quarter (2009-2018) and including nominal effort

##Loading required Packages
```{r, eval=TRUE}
require(spict)
require(gdata)
```

##Required data for Ex.2
Landings came from Spain and Morocco are in tons by quarter (2009-2018)
Biomass index: standardized CPUE (k/h)from Spain VMS data by quarter (2009-2018) 
```{r, eval=TRUE}
dir()
dat2plus <- read.xls("sbr_quarterlydata.xls")
```

##"inp" for the SPiCT model. 
```{r, eval=TRUE}
inp3plus <- list(timeC = dat2plus[,1],
            obsC = dat2plus[,2],
            timeI = dat2plus[,1]+ 0.125,  
            obsI = dat2plus[,3],
            timeE = dat2plus[,1],  
            obsE = dat2plus[,5])
inp3plus <- check.inp(inp3plus)
```

##Plotting data inputs
```{r}
plotspict.data(inp3plus)
plotspict.ci(inp3plus) ##Advanced plots
```
Error in base.checks(inp$obsE, inp$timeE, inp$stdevfacE, "E") : Time and observation vector do not match in length for  E series (meaning?)

##Fitting the model
```{r}
fit3plus <- fit.spict(inp3plus)
```

##Summary
```{r}
capture.output(summary(fit3plus))
```

#Plotting outputs
```{r}
plot(fit3plus)
```
Really uncertain but converge...short data series and seems like doesn´t get any improve including the effort :-(

