---
title: "SPiCTing with Strait of Gibraltar blackspot seabream target fishery data (Spain and Morocco) - Exercise 1+"
author: "J. Gil"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data used
A long time-series of species catch  and nominal effort by year (1983-2018) and including a short series of biomass index (2009-2018)

##Loading required Packages
```{r, eval=TRUE}
require(spict)
require(gdata)
```

##Required data for Ex.1
Landings came from Spain and Morocco are in tons by year (1983-2018)
Effort instead of CPUE (nominal effort: nº sales, Spanish ones corrected from VMS data since 1997)
```{r, eval=TRUE}
dir()
dat1plus <- read.xls("sbr_yearlydata.xls")
```

##"inp" for the SPiCT model. 
```{r, eval=TRUE}
inp1plus <- list(timeC = dat1plus[,1],
            obsC = dat1plus[,2],
            timeI = dat1plus[,1]+ 0.5,  
            obsI = dat1plus[,3],
            timeE = dat1plus[,1], 
            obsE = dat1plus[,5])
inp1plus <- check.inp(inp1plus)
```

##Plotting data inputs
```{r}
plotspict.data(inp1plus)
plotspict.ci(inp1plus) ##Advanced plots
```
Advanced plots error:
"Error in base.checks(inp$obsE, inp$timeE, inp$stdevfacE, "E") : Time and observation vector do not match in length for E series"

##Fitting the model
```{r}
fit1plus <- fit.spict(inp1plus)
```

##Summary
```{r}
capture.output(summary(fit1plus))
```

#Plotting outputs
```{r}
plot(fit1plus)
```
Quite uncertain but converge

##Alternate model: fixing to Schaefer model
```{r}
inp2plus <- inp1plus
inp2plus$phases$logn <- -1
inp2plus$ini$logn <- log(2)
inp2plus$priors$logn <- c(0,0,0)

fit2plus <- fit.spict(inp2plus)

plot(fit2plus)

capture.output(summary(fit2plus))
```
Not a huge improve :-(

# Checklist for the acceptance of a SPiCT assessment 

- 1: Convergence 
```{r}
fit1plus$opt$convergence 
fit2plus$opt$convergence 
```
Both models converge :-)

- 2: All variance parameters are finite 
```{r}
all(is.finite(fit1plus$sd))  
all(is.finite(fit2plus$sd))  
```
TRUE in both models :-)

- 3: No violation of model assumptions 
```{r}
res1plus <- calc.osa.resid(fit1plus)
plotspict.diagnostic(res1plus)
res2plus <- calc.osa.resid(fit2plus)
plotspict.diagnostic(res2plus)
```
Shapiro p-value significant in catches from Schaefer model (normality of the residuals)!! 
Model 1+ all in green :-)

- 4: Consistent patterns in the Retrospective Analysis
```{r}
rep1plus=retro(fit1plus, nretroyear=5)
plotspict.retro(rep1plus)
rep2plus=retro(fit2plus, nretroyear=5)
plotspict.retro(rep2plus)
```
Last assessment quite different in model 1+ :-( 
Also in model 2+, mostly in absolute values (B and F)

- 5. Realistic production curve
```{r}
calc.bmsyk(fit1plus)
calc.bmsyk(fit2plus) # ALTERNATE Schaefer fix (0.5)
```
0.7408798 in model 1 :-(

- 6. High assessment uncertainty can indicate a lack of contrast in the input data or violation of
the ecological model assumptions
```{r}
calc.om(fit1plus)
calc.om(fit2plus)
```
F/Fmsy confidence intervals order magnitude= 3  in model 1+ :-( 
B/Bmsy confidence intervals order magnitude= 2  in model 2+ :-( 

- 7. Initial values do not influence the parameter estimates
```{r}
check.ini(inp1plus, ntrials=33)
check.ini(inp2plus, ntrials=33)
```
3 trials (from 33) didn´t converge in model 1+. Distance close to zero (0.01 and 0.02 in certain trials) in $resmat
All trials (from 33)  converge in model 2+. Distance in $resmat is 0.10 in one trial

# Extract values in TABLE format
Preparing TABLE format
```{r, include=FALSE}
xtab<-function(x,caption='Table X.', file=stdout(), width='"100%"', cornername='', dec=rep(1,ncol(x))){
  nc<-ncol(x)
  lin<-paste('<table width=',width,'>', sep='')
  lin<-c(lin,sub('$','</td></tr>',sub('\\. |\\.$','.</b> ',
                                      sub('^', paste('<tr><td colspan=',nc+1,'><b>',sep=''), caption))))
  hr<-paste('<tr><td colspan=',nc+1,'><hr noshade></td></tr>', sep='')
  lin<-c(lin,hr)
  cnames<-colnames(x)
  cnames<-paste(sub('$','</b></td>',sub('^','<td align=right><b>',cnames)), collapse='\t')
  lin<-c(lin,paste('<tr>',paste('<td align=left><b>',cornername,'</b></td>',sep=''),cnames,'</tr>'))
  lin<-c(lin,hr)
  rnames<-sub('$','</b></td>',sub('^','<tr> <td align=left><b>',rownames(x)))
  #x<-sapply(1:ncol(x),function(i)sub('NA','  ',format(round(x[,i],dec[i]))))
  x<-sapply(1:ncol(x),function(i)sub('NA','  ',formatC(round(x[,i],dec[i]),digits=dec[i], format='f')))
  for(i in 1:nrow(x)){
    thisline<-paste(rnames[i],paste(sub('$','</td>',sub('^','<td align=right>',x[i,])), collapse='\t'),'</tr>', sep='')
    lin<-c(lin,thisline)
  }
  lin<-c(lin,hr)
  lin<-c(lin,'</table><br>\n')
  writeLines(lin,con=file)
}
```

Then choose the model and its values can be extracted with these tables:
```{r}
Tab1 <- sumspict.parest(fit1plus); Tab1
Tab2 <- sumspict.srefpoints(fit1plus);Tab2
Tab3 <- sumspict.states(fit1plus);Tab3
Tab4 <- sumspict.predictions(fit1plus);Tab4
Tab1_Sfix <- sumspict.parest(fit2plus); Tab1_Sfix
Tab2_Sfix <- sumspict.srefpoints(fit2plus);Tab2_Sfix
Tab3_Sfix <- sumspict.states(fit2plus);Tab3_Sfix
Tab4_Sfix <- sumspict.predictions(fit2plus);Tab4_Sfix
```

