---
title: "SPiCTing with Strait of Gibraltar blackspot seabream target fishery data (Spain and Morocco) - Exercise 1"
author: "J. Gil"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data used
A long time-series of species catch  and nominal effort by year (1983-2018).

##Loading required Packages
```{r, eval=TRUE}
require(spict)
require(gdata)
```

##Required data for Ex.1
Landings came from Spain and Morocco are in tons by year (1983-2018)
Effort instead of CPUE (nominal effort: nº sales, Spanish ones corrected from VMS data since 1997)
```{r, eval=TRUE}
dir()
dat1 <- read.xls("sbr_yearlydata.xls")
```

##"inp" for the SPiCT model. 
```{r, eval=TRUE}
inp1 <- list(timeC = dat1[,1],
            obsC = dat1[,2],
            timeE = dat1[,1], 
            obsE = dat1[,5])
inp1 <- check.inp(inp1)
```

##Plotting data inputs
```{r}
plotspict.data(inp1)
plotspict.ci(inp1) ##Advanced plots
```
Advanced plots are the same because there´s no Index, rigth?

##Fitting the model
```{r}
fit1 <- fit.spict(inp1)
```

##Summary
```{r}
capture.output(summary(fit1))
```

#Plotting outputs
```{r}
plot(fit1)
```
Quite uncertain but converge

##Alternate model: fixing to Schaefer model
```{r}
inp2 <- inp1
inp2$phases$logn <- -1
inp2$ini$logn <- log(2)
inp2$priors$logn <- c(0,0,0)

fit2 <- fit.spict(inp2)

plot(fit2)

capture.output(summary(fit2))
```
No improving at all :-(

# Checklist for the acceptance of a SPiCT assessment 

- 1: Convergence 
```{r}
fit1$opt$convergence 
fit2$opt$convergence 
```
Both models converge :-)

- 2: All variance parameters are finite 
```{r}
all(is.finite(fit1$sd))  
all(is.finite(fit2$sd))  
```
TRUE in both models :-)

- 3: No violation of model assumptions 
```{r}
res1 <- calc.osa.resid(fit1)
plotspict.diagnostic(res1)
res2 <- calc.osa.resid(fit2)
plotspict.diagnostic(res2)
```
Shapiro p-value significant in catches from both models (normality of the residuals)!!

- 4: Consistent patterns in the Retrospective Analysis
```{r}
rep1=retro(fit1, nretroyear=5)
plotspict.retro(rep1)
rep2=retro(fit2, nretroyear=5)
plotspict.retro(rep2)
```
Not so bad :-)

- 5. Realistic production curve
```{r}
calc.bmsyk(fit1)
calc.bmsyk(fit2) # ALTERNATE Schaefer fix (0.5)
```
0.7370801 in model 1 :-(

- 6. High assessment uncertainty can indicate a lack of contrast in the input data or violation of
the ecological model assumptions
```{r}
calc.om(fit1)
calc.om(fit2)
```
B/Bmsy confidence intervals order magnitude= 2 in model 2 :-( 

- 7. Initial values do not influence the parameter estimates
```{r}
check.ini(inp1, ntrials=33)
check.ini(inp2, ntrials=33)
```
4 trials (from 33) didn´t converge in model 1. Distance close to zero in $resmat
4 trials (from 33) didn´t converge in model 2. Distance in $resmat is 0.22 in one trial

# Extract values in TABLE format
Preparing TABLE format
```{r, include=FALSE}
xtab<-function(x,caption='Table X.', file=stdout(), width='"100%"', cornername='', dec=rep(1,ncol(x))){
  nc<-ncol(x)
  lin<-paste('<table width=',width,'>', sep='')
  lin<-c(lin,sub('$','</td></tr>',sub('\\. |\\.$','.</b> ',
                                      sub('^', paste('<tr><td colspan=',nc+1,'><b>',sep=''), caption))))
  hr<-paste('<tr><td colspan=',nc+1,'><hr noshade></td></tr>', sep='')
  lin<-c(lin,hr)
  cnames<-colnames(x)
  cnames<-paste(sub('$','</b></td>',sub('^','<td align=right><b>',cnames)), collapse='\t')
  lin<-c(lin,paste('<tr>',paste('<td align=left><b>',cornername,'</b></td>',sep=''),cnames,'</tr>'))
  lin<-c(lin,hr)
  rnames<-sub('$','</b></td>',sub('^','<tr> <td align=left><b>',rownames(x)))
  #x<-sapply(1:ncol(x),function(i)sub('NA','  ',format(round(x[,i],dec[i]))))
  x<-sapply(1:ncol(x),function(i)sub('NA','  ',formatC(round(x[,i],dec[i]),digits=dec[i], format='f')))
  for(i in 1:nrow(x)){
    thisline<-paste(rnames[i],paste(sub('$','</td>',sub('^','<td align=right>',x[i,])), collapse='\t'),'</tr>', sep='')
    lin<-c(lin,thisline)
  }
  lin<-c(lin,hr)
  lin<-c(lin,'</table><br>\n')
  writeLines(lin,con=file)
}
```

Then choose the model and its values can be extracted with these tables:
```{r}
Tab1 <- sumspict.parest(fit1); Tab1
Tab2 <- sumspict.srefpoints(fit1);Tab2
Tab3 <- sumspict.states(fit1);Tab3
Tab4 <- sumspict.predictions(fit1);Tab4
Tab1_Sfix <- sumspict.parest(fit2); Tab1_Sfix
Tab2_Sfix <- sumspict.srefpoints(fit2);Tab2_Sfix
Tab3_Sfix <- sumspict.states(fit2);Tab3_Sfix
Tab4_Sfix <- sumspict.predictions(fit2);Tab4_Sfix
```

