---
title: "SPiCTing with Strait of Gibraltar blackspot seabream target fishery data (Spain and Morocco) - Exercise 3"
author: "J. Gil"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data used
Mixing data by year and quarters from Ex.1 and Ex.2

##Loading required Packages
```{r, eval=TRUE}
require(spict)
require(gdata)
```

##Required data for Ex.3
Landings came from Spain and Morocco are in tons by year (1983-2008) and by quarter (2009-2018)
Biomass index (2009-2018)  and nominal effort (nº sales, Spanish ones corrected from VMS data since 1997) 
```{r, eval=TRUE}
dir()
dat3 <- read.xls("sbr_YandQ.xls")
```

##"inp" for the SPiCT model. 
```{r, eval=TRUE}
inp5 <- list(timeC = dat3[,1],
            obsC = dat3[,2],
            timeI = dat3[,1]+ 0.125,  
            obsI = dat3[,3],
            timeE = dat3[,1],  
            obsE = dat3[,5])
inp5 <- check.inp(inp5)
```

##Plotting data inputs
```{r}
plotspict.data(inp5)
```

##Fitting the model
```{r}
fit5 <- fit.spict(inp5)
```

##Summary
```{r}
capture.output(summary(fit5))
```

#Plotting outputs
```{r}
plot(fit5)
```
Sometimes you required change the type of the reference point to deterministic, because there´s not biomass relative to Bmsy (NA) Figure 
```{r}
inp5$msytype <- "d"

fit5 <- fit.spict(inp5)

plot(fit5)

capture.output(summary(fit5))
```
Really weird...something wrong, can we mix year and quarter data series?

##Alternate model: fixing to Schaefer model
```{r}
inp6 <- inp5
inp6$phases$logn <- -1
inp6$ini$logn <- log(2)
inp6$priors$logn <- c(0,0,0)
inp6$msytype <- "d"

fit6 <- fit.spict(inp6)

plot(fit6)

capture.output(summary(fit6))
```
Again, really weird :-(
